---
title: "Key Inequality Indicators"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    theme: paper
    source_code: embed
    vertical_layout: scroll
runtime: shiny
---

```{r setup, include=FALSE}
library("here")
library("data.table")
library("ggplot2")
library("plotly")
library("flexdashboard")  
library("tidyverse")
library("pak")
library("rlang")
library("joyn")
```



```{r setup-run-scripts, include=FALSE}

lapply(
  list.files(here::here("R"), full.names = T), 
  function(flist){source(flist, echo = F, verbose = F)}
)


```






Deep-dive on specific country indicators over time {data-icon="fa-arrow-trend-up"}
===========================================================


Sidebar {.sidebar data-width=250}
-----------------------------------------------------------------------
```{r}
# For 
selectInput( # plot_over_time_ineq
        "country_a",
        label = "Select country",
        choices = pip_countries,
        selected = c("South Africa"),
        multiple = FALSE
      )

selectInput(  
        "indicator1_a",
        label = "Select first inequality indicator",
        choices = c(
      "Mean", 
      "Median", 
      "Mld", 
      "Gini", 
      "Decile1", 
      "Decile2", 
      "Decile3", 
      "Decile4", 
      "Decile5", 
      "Decile6", 
      "Decile7", 
      "Decile8", 
      "Decile9", 
      "Decile10",
      "Top10_bottom40", 
      "Top20_bottom20"
    ),
        selected = "Gini",
        multiple = FALSE
      )
selectInput(  
        "indicator2_a",
        label = "Select second inequality indicator",
        choices = c(
      "Mean", 
      "Median", 
      "Mld", 
      "Gini", 
      "Decile1", 
      "Decile2", 
      "Decile3", 
      "Decile4", 
      "Decile5", 
      "Decile6", 
      "Decile7", 
      "Decile8", 
      "Decile9", 
      "Decile10",
      "Top10_bottom40", 
      "Top20_bottom20"
    ),
        selected = "Mld",
        multiple = FALSE
      )


```


Row
-----------------------------------------------------------------------

### Gini

```{r}

# dt1 <- reactive({
#   dt1 <- dt_pip[
#     country_name == input$country_a, 
#     .(year, gini, median, top10_bottom40)
#   ]
# })

renderValueBox(
  valueBox(
    paste("Gini in", dt_pip[country_name == input$country_a, .(year, gini, median, top10_bottom40)]$year |> last() |> round(2)),
  dt_pip[
    country_name == input$country_a, 
    .(year, gini, median, top10_bottom40)
  ]$gini |> last() |> round(2)
))
#valueBox(0.63, icon = "fa-comments")

#gauge(91, min = 0, max = 100, symbol = '%', gaugeSectors(
#  success = c(80, 100), warning = c(40, 79), danger = c(0, 39)
#))
```

### Median 

```{r}
renderValueBox(
  valueBox(
    paste("Median in", dt_pip[country_name == input$country_a, .(year, gini, median, top10_bottom40)]$year |> last() |> round(2)),
  dt_pip[
    country_name == input$country_a, 
    .(year, gini, median, top10_bottom40)
  ]$median |> last() |> round(2)
))
```

### Top 10 Bottom 40 Ratio

```{r}
renderValueBox(
  valueBox(
    paste("Top 10 Bottom 40 ratio in", dt_pip[country_name == input$country_a, .(year, gini, median, top10_bottom40)]$year |> last() |> round(2)),
  dt_pip[
    country_name == input$country_a, 
    .(year, gini, median, top10_bottom40)
  ]$top10_bottom40 |> last() |> round(2)
))
```

### Year, Country

```{r}
renderText(
  print(
    paste("For", dt_pip[country_name == input$country_a, .(year, gini, median, top10_bottom40)]$year |> last() |> round(2)), "in", dt_pip[country_name == input$country_a, .(year)]$year |> last() |> round(2))
  )
```


Row
-----------------------------------------------------------------------

### Changing inequality rankings - first indicator

```{r}
renderPlot(
  plot_over_time_ineq(
  dt               = dt_pip, 
  country_selected = input$country_a, 
  ineq_indicator   = input$indicator1_a
)
)


```


Row
-----------------------------------------------------------------------

### Changing inequality rankings - second indicator

```{r}
renderPlot(
plot_over_time_ineq(
  dt               = dt_pip, 
  country_selected = input$country_a, 
  ineq_indicator   = input$indicator2_a
)
)

```



Changing inequality rankings {data-icon="fa-ranking-star"}
===========================================================

Sidebar {.sidebar data-width=350}
-----------------------------------------------------------------------
```{r}
# For 
selectInput( # plot_over_time_ineq
        "country_b",
        label = "Select country",
        choices = pip_countries,
        selected = "",
        multiple = FALSE
      )

selectInput(  
        "indicator1_b",
        label = "Select first inequality indicator",
        choices = c(
      "Mean", 
      "Median", 
      "Mld", 
      "Gini", 
      "Decile1", 
      "Decile2", 
      "Decile3", 
      "Decile4", 
      "Decile5", 
      "Decile6", 
      "Decile7", 
      "Decile8", 
      "Decile9", 
      "Decile10",
      "Top10_bottom40", 
      "Top20_bottom20"
    ),
        selected = "Gini",
        multiple = FALSE
      )
selectInput(  
        "indicator2_b",
        label = "Select second inequality indicator",
        choices = c(
      "Mean", 
      "Median", 
      "Mld", 
      "Gini", 
      "Decile1", 
      "Decile2", 
      "Decile3", 
      "Decile4", 
      "Decile5", 
      "Decile6", 
      "Decile7", 
      "Decile8", 
      "Decile9", 
      "Decile10",
      "Top10_bottom40", 
      "Top20_bottom20"
    ),
        selected = "Gini",
        multiple = FALSE
      )




#selectInput( # plot_indicator_ranking_change
#        "region",
#        label = "Select region",
#        choices = dt_pip$region_code |> unique(),
#        selected = "SSA",
#        multiple = FALSE
#      )


#selectInput("start_yr", # 
#            label    = "Start of five year window",
#            choices  = dt_pip$year |> unique(), 
#            selected = 2000, 
#            multiple = FALSE)

#radioButtons("chosen_yrs", 
#              label = "Selected years",
#              choices = 
#                c("Latest",
#                  "All"),
#              selected = "Latest"
#              )




# 
# sliderInput("pc", 
#             label = "Percentile",
#             min = 10, max = 100, value = 50, step = 10)
# 
# 
# sliderInput("hc", 
#             label = "Decile within percentile",
#             min = 1, max = 10, value = 10, step = 1)
# 
# selectInput(
#         "ctr",
#         label = "Select countries to highlight",
#         choices = NULL,
#         selected = "",
#         multiple = TRUE
#       )



```


Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Changing inequality rankings - first indicator 

```{r}



```


-----------------------------------------------------------------------

### Changing inequality rankings - second indicator

```{r}

```














Compare indicators across countries {data-icon="fa-table"}
===========================================================


Sidebar {.sidebar data-width=350}
-----------------------------------------------------------------------

```{r}
# For 
selectInput( # make_table_across_countries
        "countries_c",
        label = "Select countries",
        choices = pip_countries,
        selected = c("South Africa", "Colombia", "United States"),
        multiple = TRUE
      )

selectInput(
        "indicators_c",
        label = "Select inequality indicators",
        choices = c(
      "Mean", 
      "Median", 
      "Mld", 
      "Gini", 
      "Decile1", 
      "Decile2", 
      "Decile3", 
      "Decile4", 
      "Decile5", 
      "Decile6", 
      "Decile7", 
      "Decile8", 
      "Decile9", 
      "Decile10",
      "Top10_bottom40", 
      "Top20_bottom20"
    ),
        selected = "Gini",
        multiple = TRUE
      )



radioButtons("chosen_yrs_c", 
              label = "Selected years",
              choices = 
                c("Latest",
                  "All"),
              selected = "Latest"
              )


```


Row
-----------------------------------------------------------------------

### Table of inequality indicators

```{r}


renderPlotly(
  make_table_across_countries(
    dt               = dt_pip, 
    country_name_vec = input$countries_c, 
    ineq_indicators  = input$indicators_c, 
    years_selected   = input$chosen_yrs_c
  )
)

```


